Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}

### 2020-12-08 article.Rnw
update draft and create Rnw structure

\documentclass[article]{jss}
\usepackage{amsmath}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Anonymous Giraffe\\Universit\"at Innsbruck \And 
        Second Author\\Plus Affiliation}
\title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Anonymous Giraffe, Second Author} %% comma-separated
\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using \texttt{scoringutils} in R} %% without formatting
\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Forecasts play a role in many scientific fields from finance to meteorology and epidemiology. With the emergence of Covid-19 the role of forecasting to inform public policy has again attracted increased attention. In this paper we introduce a complete framework for evaluating different types of forecasts using the R package scoringutils. We discuss different appropriate evaluation metrics and apply the framework to XXXXX. 
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Nikos Bosse\\
  Centre for Mathematical Modelling of Infectious Diseases\\
%%  Faculty of Economics and Statistics\\
  London School of Hygiene and Tropical Medicine\\
%%  6020 Innsbruck, Austria\\
  E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
  %% URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section[introduction]{Introduction}
Good forecasts are of great interest to decision makers in almost every field. An integral part of assessing and improving their usefulness is forecast evaluation. For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts [some citations]. Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial. We therefore present the \pkg{scoringutils} pacakge. The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts. In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK.

@@

Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007]. For an ideal forecast, we therefore have 
$$ P_t = F_t, $$
where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution. As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly. [CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness. Calibration refers to the statistical consistency between the predictive distribution and the observations. A well calibrated forecast does not systematically differ deviate from the observed values. For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007]. Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e. how precise the forecasts are. The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration. Take for example the task of predicting rainfall in a city like London. A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time. On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day. Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again. DO I WANT THAT FIGURE?

@@

\section[run-evaluation]{Running an automated evaluation}
A full evaluation of all forecasts based on observed values can be performed using the function \code{eval_forecasts}. The function automatically recognises the prediction type and format and applies the appropriate scoring metrics. Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.

@@

\begin{CodeInput}
# install and load data pacakge from external repository
remotes::install_github("sbfnk/covid19.forecasts.uk")
library(covid19.forecasts.uk)

# load truth data
data(covid_uk_data)

# load forecasts
data(uk_forecasts)
\end{CodeInput}

ADD OUTPUT HERE TO SHOW THE FIRST COUPLE OF ROWS (MUST BE DONE WITH SWEAVE PROBABLY)

@@

\begin{CodeInput}
combined <- dplyr::inner_join(uk_forecasts %>%
                                dplyr::rename(prediction = value), 
                              covid_uk_data %>%
                                dplyr::rename(true_value = value))
\end{CodeInput}
MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY

The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call 
\begin{CodeInput}
additional_observations <- covid_uk_data %>%
  dplyr::rename(true_value = value) %>%
  dplyr::filter(value_date <= "2020-06-22", 
                value_date > "2020-06-01", 
                value_type %in% unique(example_subset$value_type))

scoringutils::plot_predictions(example_subset, 
                               additional_observations,
                               x = "value_date",
                               facet_formula = geography ~ value_type)
\end{CodeInput}
MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.

An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
\begin{CodeInput}
library(scoringutils)
scores <- eval_forecasts(combined, 
                         summarise_by = c("model", "value_type"))

scores
\end{CodeInput}

Pairwise comparisons between models [CITATION] can be made using 
\begin{CodeInput}
pairwise <- pairwise_comparison(scores)
\end{CodeInput}
This will be explained in more detail in section [visualisation and interpretation of the evaluation].

@@

\section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}

@@

For quantile forecasts, we propose the following metric to assess bias: 
\begin{align*}
  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t\}) \mathbb{1}( x_t \leq q_{t, 0.5}) \\
  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbb{1}( x_t \geq q_{t, 0.5}),
\end{align*}
where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$. They represent our belief about what the true value $x_t$ will be. For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$. $\mathbb{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise. In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution. If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value. If the true value is exactly the median, both terms cancel out and $B_t$ is zero. For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts. For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score. These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.

@@

\section[visualisation]{Visualisation and interpretation of evaluation results}

@@

>>>>

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{amsmath, amsfonts}
\usepackage{bbm}% to get \mathbbmm{1}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE, prefix.string=plots/plot}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Nikos I. Bosse\\London School of Hygiene and Tropical Medicine
   \And Second Author\\Plus Affiliation}
\Plainauthor{Nikos I. Bosse, Second Author}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}
\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using scoringutils in R}
\Shorttitle{Evaluating Forecasts with \pkg{scoringutils} in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{
  Forecasts play a role in many scientific fields from finance to meteorology and epidemiology. With the emergence of Covid-19 the role of forecasting to inform public policy has again attracted increased attention. In this paper we introduce a complete framework for evaluating different types of forecasts using the R package scoringutils. We discuss different appropriate evaluation metrics and apply the framework to XXXXX. 
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Nikos Bosse\\
  Centre for Mathematical Modelling of Infectious Diseases\\
  London School of Hygiene and Tropical Medicine\\
  Keppel Street \\ 
  London WC1E 7HT \\
  E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
  % URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}
\SweaveOpts{concordance=FALSE}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction]{Introduction}
Good forecasts are of great interest to decision makers in almost every field. An integral part of assessing and improving their usefulness is forecast evaluation. For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts [some citations]. Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial. We therefore present the \pkg{scoringutils} pacakge. The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts. In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK.

@@

Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007]. For an ideal forecast, we therefore have 
%
$$ P_t = F_t, $$
%
where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution. As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly. [CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness. Calibration refers to the statistical consistency between the predictive distribution and the observations. A well calibrated forecast does not systematically differ deviate from the observed values. For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007]. Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e. how precise the forecasts are. The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration. Take for example the task of predicting rainfall in a city like London. A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time. On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day. Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again. DO I WANT THAT FIGURE?

@@

% ==============================================================================
\section[run-evaluation]{Running an automated evaluation}

A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}. The function automatically recognises the prediction type and format and applies the appropriate scoring metrics. Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.

@@

<<load-data>>=
# install and load data pacakge from external repository
# remotes::install_github("sbfnk/covid19.forecasts.uk")
library(covid19.forecasts.uk)

# load truth data
data(covid_uk_data)
head(covid_uk_data)

# load forecasts
data(uk_forecasts)
head(uk_forecasts)
@

@@

<<modify-data>>=
combined <- dplyr::inner_join(uk_forecasts %>%
                                dplyr::rename(prediction = value), 
                              covid_uk_data %>%
                                dplyr::rename(true_value = value))
@

MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY

The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call

<<show-forecasts, eval=FALSE>>=
example_subset <- combined %>%
  dplyr::filter(model == "SIRCOVID", 
                creation_date == "2020-06-22")

additional_observations <- covid_uk_data %>%
  dplyr::rename(true_value = value) %>%
  dplyr::filter(value_date <= "2020-06-22", 
                value_date > "2020-06-01", 
                value_type %in% unique(example_subset$value_type))

scoringutils::plot_predictions(example_subset, 
                               additional_observations,
                               x = "value_date",
                               facet_formula = geography ~ value_type)
@

The output is shown in Figure \ref{fig:forecast-visualisation}.

MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.

\begin{figure}[h]
\centering
<<show-forecasts, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
example_subset <- combined %>%
  dplyr::filter(model == "SIRCOVID", 
                creation_date == "2020-06-22")

additional_observations <- covid_uk_data %>%
  dplyr::rename(true_value = value) %>%
  dplyr::filter(value_date <= "2020-06-22", 
                value_date > "2020-06-01", 
                value_type %in% unique(example_subset$value_type))

scoringutils::plot_predictions(example_subset, 
                               additional_observations,
                               x = "value_date",
                               facet_formula = geography ~ value_type)
@
\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
\end{figure}

An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
<<scoring>>=
library(scoringutils)
scores <- eval_forecasts(combined, 
                         summarise_by = c("model", "value_type"))

dplyr::glimpse(scores)
@

Pairwise comparisons between models [CITATION] can be made using 
<<pairwise-comparison>>=
pairwise <- pairwise_comparison(scores)
@

This will be explained in more detail in section [visualisation and interpretation of the evaluation].

@@

% ==============================================================================
\section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}

@@

For quantile forecasts, we propose the following metric to assess bias: 
\begin{align*}
  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
\end{align*}
where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$. They represent our belief about what the true value $x_t$ will be. For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$. $\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise. In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution. If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value. If the true value is exactly the median, both terms cancel out and $B_t$ is zero. For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts. For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score. These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.

@@

\section{Visualisation and interpretation of evaluation results}

A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:

\begin{figure}[h]
\centering
<<score-table, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
scoringutils::score_table(scores, y = "model", facet_formula = ~ value_type)
@
\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
\end{figure}

If all we care about is which model performed best, the pairwise comparison is most suitable. This approach compares all possible pairs of models based on an overlapping set of forecasts.

\begin{figure}[h]
\centering
<<pairwise-plot, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
scoringutils::plot_pairwise_comparison(pairwise)
@
\caption{\label{fig:pairwise-comparison} XXXX}
\end{figure}

@@

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

% The results in this paper were obtained using
% \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
% \pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

% All acknowledgments (note the AE spelling) should be collected in this
% unnumbered section before the references. It may contain the usual information
% about funding and feedback from colleagues/reviewers/etc. Furthermore,
% information such as relative contributions of the authors may be added here
% (if any).

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{Appendix section} \label{app:technical}

\end{appendix}

%% -----------------------------------------------------------------------------
